---
title: "Food Allergens Prediction"
output: html_document
---

## Install and load packages

```{r message=FALSE, warning=FALSE}
# install.packages(c('caTools', 'caret', 'vip','randomForest'))
library(caTools)
library(caret)
library(dplyr)
library(randomForest)
library(vip)
```

## Import the dataset

```{r}
food <- read.csv("C:/Users/Admin/Documents/GitHub/Food_Allergens/food_ingredients_and_allergens.csv")
```

## Have a look at the dataframe

```{r}
head(food)
colnames(food)
str(food)
```

## Handle duplicate rows

```{r}
food <- distinct(food)
```

## Encode the target variable - Prediction

```{r}
food$Prediction = factor(food$Prediction,
                         levels = c('Contains', 'Does not contain'),
                         labels = c(0,1))
```

## Add factors for variables that are factors

```{r}

food$Allergens <- as.factor(food$Allergens)
food$Food.Product <- as.factor(food$Food.Product)
food$Main.Ingredient <- as.factor(food$Main.Ingredient)
food$Fat.Oil <- as.factor(food$Fat.Oil)
food$Seasoning <- as.factor(food$Seasoning)
food$Sweetener <- as.factor(food$Sweetener)
str(food)
```

## Handle missing values

```{r}
colSums(is.na(food))
food <- na.omit(food)
glimpse(food)
summary(food)
```
There is one missing value in the Prediction column. Since this is the target variable, I dropped one row with missing value.

Include only "Allergens", "Sweetener", "Fat.Oil", "Prediction" columns.

```{r}
food <- food %>%
  select("Allergens", "Sweetener", "Fat.Oil", "Prediction")
```

## Prepare to run the random forest

```{r}
# Split the dataset
set.seed(123)
split <- sample.split(food$Prediction, SplitRatio = 0.8)
training_set <- subset(food, split == TRUE)
test_set <- subset(food, split == FALSE)
```

```{r}
# Build the random forest
classifier <- randomForest(Prediction ~ ., data=training_set, ntree=500, proximity=TRUE)
classifier
```

```{r}
# Predict on the test set
pred <- predict(classifier, newdata = test_set[-4])
pred
```

```{r}
# Create the confusion matrix
cm <- confusionMatrix(pred, test_set$Prediction)
cm
```

The model correctly predicts 33 dishes that contain allergens, and correctly predicts 22 dishes that do not contain allergens. However, the model does not correctly predict 6 dishes that contain allergens. The Accuracy of the model is 90%.

```{r}
# Create a variable importance plot
var_importance <- vip::vip(classifier, num_features = 10)
print(var_importance)
```

Based on the above plot, we can infer the most important variable is Allergens which is the presence of ingredients might trigger allergies.